\textbf{Replies to Reviewers' Comments}

### 2023-11-25 Articles/2023-HilbertQuantum/PhysScripta/response1.tex
Comments from Francisco

\noindent First of all, we would like to thank the editorial board for the consideration of our article. Stepping back, the main line of inquiry of this paper, and our work more in general, is determining whether a particular mathematical structure is the ``correct'' one to represent a physical system. While we are convinced that no real progress can be made in the foundations of physics without establishing a precise link between mathematical models and the physical systems they are supposed to represent, the academic community does not seem set up to encourage these discussions. Since the problem sits right in between mathematics and physics, most physicists do not have enough of the technical mathematical background required, and most mathematicians do not have the physical intuition that in the end must guide the decisions.

@@

In the cover letter we specifically recommended ``not relying exclusively on reviewers who would focus primarily on the mathematical aspects'' because ``the core new idea is to look for physical justification of the mathematical structure, not simply internal mathematical consistency.'' We found, in fact, that the vast majority of physicists who read the paper concurred that having a time evolution oscillating between finite and infinite expectation values is an objective problem. On the other hand, most mathematicians either abstained (as it is a physical matter), or argued that the mathematics is consistent, and therefore there was no problem. Both reviewers, in different manners, did that, and as predicted, ``missed the main point of the paper.''

@@

The reviewer continues
\begin{quote}
	On the other hand, in Section C, the authors provide two examples of wave
	functions. From the elementary definition of an operator, we need to specify its
	domain and the rule that assigns the image of each element of its domain. In
	particular, if we had the operator $x^n$ in $L^2$, the maximal domain of definition
	is $\{\psi \in L^2 : x^n \psi \in L^2\}$. The wave function given in Equation 1 belongs to any
	of these domains, for n = 0, 1, . . . ; however, the wave function in Equation 2
	will not belong to such domains since $n \geq 1$. Henceforth, the statement that
	``expectation values will diverge for all moments'' does not make sense because
	the wave function given in Equation 1 does not belong to the domain of any
	power $x^n$ for $n \geq 1$.
\end{quote}
The reviewer points out that, mathematically, it would be more proper to say that the expectation of the observable is not defined, instead of being infinite. Therefore, we should say that the expectation value of the energy, or the position, or the number of particles, is not defined. How does this solve the physical problem? Arguably, saying the expectation of the energy of a particular state is undefined is even worse than saying that it is infinite.

@@

Changes of coordinates are a basic and fundamental notion in physics, and a pre-requisite to follow many physics arguments, including the one we make. The reviewer does not demonstrate a good understanding of the subject. Moreover, the reviewer shows no appreciation for the physical problem: they do not seem concerned that quantities like position, momentum, energy, number of particles, or any other unbounded quantities are not well defined on all states. This is exactly the kind of attitude that, as mentioned above, prevents the advancement of the mathematical foundations of physics.

@@

The reviewer concludes
\begin{quote}
A final comment. The reviewer argued above that the way that a random
variable is realized is by doing the experiment (perhaps many times) and examining
the actual outcome values. This is one very standard approach to
interpreting probability.
In quantum mechanics this works for a self-adjoint operator, or even for
a family of commuting self-adjoint operators. However there is the following
result. It is so general that it does not even require an assumption that the
experiment is repeated.
Theorem. Consider a finite-dimensional Hilbert space of dimension at least
4. Suppose that for each self-adjoint operator A there is an “experimental”
value v(A). Suppose these values satisfy at least the following two properties:
• v(A) is always an eigenvalue of A
• A commutes with B implies v(A + B) = v(A) + v(B).
Then this leads to a contradiction. For a proof see Bricmont, Goldstein,
Hemmick: From EPR-Schrodinger paradox to nonlocality based on perfect correlations.
In other words, the restriction to a commuting family is essential. But what
is the precise mathematical criterion for choosing the particular family? That
is the real problem of “physicality”.
\end{quote}
The reviewer now discusses interpretations of probability, and a theorem related to hidden variables. He claims that the real problem of ``physicality'' lies in understanding these issues. This makes it clear that, in the eye of the reviewer, we are not working on ``real'' issues.

>>>>

\noindent First of all, we would like to thank the editorial board for the consideration of our article. Stepping back, the main line of inquiry of this paper, and our work more in general, is determining whether a particular mathematical structure is the ``correct'' one to represent a physical system. While we are convinced that no real progress can be made in the foundations of physics without establishing a precise link between mathematical models and the physical systems they are supposed to represent, the academic community does not seem set up to encourage these discussions. Since the problem sits right in between mathematics and physics, and given that the fields have separate training, conferences, journals, culture, etc.,  most physicists do not have enough of the technical mathematical background required, and most mathematicians do not have the physical intuition that in the end must guide the decisions.

@@

In the cover letter we specifically recommended ``not relying exclusively on reviewers who would focus primarily on the mathematical aspects'' because ``the core new idea is to look for physical justification of the mathematical structure, not simply internal mathematical consistency.'' We found, in fact, that the vast majority of physicists who read the paper concurred that having a time evolution oscillating between finite and infinite expectation values is an objective problem. On the other hand, most mathematicians either abstained (as it is a physical matter), or argued that the mathematics is consistent, and therefore there was no problem. Both reviewers focused mainly on the math, and as predicted, ``missed the main point of the paper.''

@@

The reviewer continues
\begin{quote}
	On the other hand, in Section C, the authors provide two examples of wave
	functions. From the elementary definition of an operator, we need to specify its
	domain and the rule that assigns the image of each element of its domain. In
	particular, if we had the operator $x^n$ in $L^2$, the maximal domain of definition
	is $\{\psi \in L^2 : x^n \psi \in L^2\}$. The wave function given in Equation 1 belongs to any
	of these domains, for n = 0, 1, . . . ; however, the wave function in Equation 2
	will not belong to such domains since $n \geq 1$. Henceforth, the statement that
	``expectation values will diverge for all moments'' does not make sense because
	the wave function given in Equation 1 does not belong to the domain of any
	power $x^n$ for $n \geq 1$.
\end{quote}
The reviewer points out that, mathematically, it would be more proper to say that the expectation of the observable is not defined, instead of being infinite. Therefore, we should say that the expectation value of the energy, or the position, or the number of particles, is not defined. How does this solve the physical problem? Arguably, saying the expectation of the energy of a particular state is undefined is even worse than saying that it is infinite. The carefulness in specifying the domain seems to be motivated by having all the mathematical objects being well-defined, not all the mathematical objects adequately representing the physical entities they intend to represent as is the goal of our paper.

@@

Changes of coordinates are a basic and fundamental notion in physics, and a pre-requisite to follow many physics arguments, including the one we make. The reviewer does not demonstrate a good understanding. Moreover, the reviewer shows no appreciation for the physical problem: they do not seem concerned that quantities like position, momentum, energy, number of particles, or any other unbounded quantities are not well defined on all states. This is exactly the kind of attitude that, as mentioned above, prevents the advancement of the mathematical foundations of physics.

@@

The reviewer concludes
\begin{quote}
A final comment. The reviewer argued above that the way that a random
variable is realized is by doing the experiment (perhaps many times) and examining
the actual outcome values. This is one very standard approach to
interpreting probability.
In quantum mechanics this works for a self-adjoint operator, or even for
a family of commuting self-adjoint operators. However there is the following
result. It is so general that it does not even require an assumption that the
experiment is repeated.
Theorem. Consider a finite-dimensional Hilbert space of dimension at least
4. Suppose that for each self-adjoint operator A there is an “experimental”
value v(A). Suppose these values satisfy at least the following two properties:
• v(A) is always an eigenvalue of A
• A commutes with B implies v(A + B) = v(A) + v(B).
Then this leads to a contradiction. For a proof see Bricmont, Goldstein,
Hemmick: From EPR-Schrodinger paradox to nonlocality based on perfect correlations.
In other words, the restriction to a commuting family is essential. But what
is the precise mathematical criterion for choosing the particular family? That
is the real problem of “physicality”.
\end{quote}
The reviewer now discusses interpretations of probability, and a theorem related to hidden variables. They claim that the real problem of ``physicality'' lies in understanding these issues. This makes it clear that, in the eye of the reviewer, we are not working on ``real'' issues.
